# Notes

This document lists UCI datasets that are potentially useful for regression projects.  
For each dataset below, I include:

- the UCI webpage link,
- an R code chunk to import the dataset,
- basic information such as number of variables, number of observations, response variable, and topic.

Most import examples use the `{ucimlrepo}` package, which can fetch datasets directly from the UCI Machine Learning Repository from within R.

```{r}
#| warning: false
# install.packages("ucimlrepo")
library(ucimlrepo)
```

# Clean datasets

<!-- 
## Yacht Hydrodynamics

Webpage: <https://archive.ics.uci.edu/dataset/243/yacht+hydrodynamics>

```{r}
#| eval: false
yacht_hydrodynamics <- fetch_ucirepo(id = 243)
yacht_hydrodynamics_data <- yacht_hydrodynamics$data$original

dim(yacht_hydrodynamics_data)
head(yacht_hydrodynamics_data)
```

- **Number of features:** 6
- **Number of targets:** 1
- **Number of observations:** 308
- **Response variable:** residuary resistance per unit weight of displacement
- **Topic:** yacht design and hydrodynamic resistance -->

# Datasets that need a little preprocessing

<!-- ## Bank Marketing

Webpage: <https://archive.ics.uci.edu/dataset/222/bank+marketing>

```{r}
#| eval: false
bank_marketing <- fetch_ucirepo(id = 222)
bank_marketing_data <- bank_marketing$data$original

dim(bank_marketing_data)
head(bank_marketing_data)
```

- **Number of features:** 16
- **Number of targets:** 1
- **Number of observations:** 45,211
- **Response variable:** official target is `y`; for regression, users sometimes instead model a quantitative field such as `balance`
- **Topic:** direct marketing campaigns by a Portuguese banking institution
- **Preprocessing needed:** remove or carefully handle the campaign outcome variable `y` if it is not being modeled
- **Preprocessing needed:** encode categorical fields such as job, marital status, education, contact type, month, and poutcome
- **Preprocessing needed:** check for skewness in `balance`, `duration`, and campaign-count variables and transform if needed -->

<!-- ## Online Retail

Webpage: <https://archive.ics.uci.edu/dataset/352/online+retail>

```{r}
#| eval: false
online_retail <- fetch_ucirepo(id = 352)
online_retail_data <- online_retail$data$original

dim(online_retail_data)
head(online_retail_data)
```

- **Number of variables:** 8
- **Number of observations:** 541,909
- **Response variable:** a quantitative choice is `UnitPrice`
- **Topic:** online retail transactions
- **Preprocessing needed:** parse `InvoiceDate` as a date-time variable
- **Preprocessing needed:** remove cancellations and returns, typically identified by invoice codes beginning with `C` or by negative quantities
- **Preprocessing needed:** decide how to handle missing `CustomerID`
- **Preprocessing needed:** encode `Country` and engineer transaction-level or customer-level features if modeling beyond raw line items -->

<!-- ## Car Evaluation

Webpage: <https://archive.ics.uci.edu/dataset/19/car+evaluation>

```{r}
#| eval: false
car_evaluation <- fetch_ucirepo(id = 19)
car_evaluation_data <- car_evaluation$data$original

dim(car_evaluation_data)
head(car_evaluation_data)
```

- **Number of features:** 6
- **Number of targets:** 1
- **Number of observations:** 1,728
- **Response variable:** this is mainly a classification dataset; if you use it for regression, you would need to create an ordinal or numeric response yourself
- **Topic:** car acceptability evaluation
- **Preprocessing needed:** create a numeric or ordinal response from the categorical acceptability label
- **Preprocessing needed:** encode all predictors, which are categorical and ordinal
- **Preprocessing needed:** preserve ordinal structure for variables such as buying price, maintenance cost, safety, luggage boot size, and number of persons -->

<!-- 
## Individual Household Electric Power Consumption

Webpage: <https://archive.ics.uci.edu/dataset/235/individual+household+electric+power+consumption>

```{r}
#| eval: false
household_power <- fetch_ucirepo(id = 235)
household_power_data <- household_power$data$original

dim(household_power_data)
head(household_power_data)
```

- **Number of features:** 9
- **Number of targets:** not clearly specified on the UCI page
- **Number of observations:** 2,075,259
- **Response variable:** a clear regression target is `Global_active_power`, the household's minute-level active power consumption
- **Topic:** household electricity consumption over time
- **Missing-data handling:** convert blank measurement fields to `NA`; the UCI page notes that about 1.25% of the measurement rows have missing values
- **Parsing / encoding:** parse `Date` and `Time`, combine them into a timestamp, and convert measurement columns from text to numeric values after import
- **Merging / reshaping:** none -->

<!-- ## Beijing Multi-Site Air Quality

Webpage: <https://archive.ics.uci.edu/dataset/501/beijing+multi+site+air+quality+data>

```{r}
#| eval: false
beijing_air_quality <- fetch_ucirepo(id = 501)
beijing_air_quality_data <- beijing_air_quality$data$original

dim(beijing_air_quality_data)
head(beijing_air_quality_data)
```

- **Number of features:** not clearly specified on the UCI page
- **Number of targets:** not clearly specified on the UCI page
- **Number of observations:** 420,768
- **Response variable:** a natural regression target is one pollutant such as `PM2.5`, `PM10`, or `O3`
- **Topic:** multi-site air-pollution and weather measurements in Beijing
- **Missing-data handling:** keep, drop, or impute rows with missing values; the UCI page states that missing data are denoted as `NA`
- **Parsing / encoding:** parse `year`, `month`, `day`, and `hour` into a timestamp
- **Merging / reshaping:** unzip and stack the multiple site files into one table, keeping `station` as a grouping variable if needed -->

<!-- ## SML2010

Webpage: <https://archive.ics.uci.edu/dataset/274/sml2010>

```{r}
#| eval: false
sml2010 <- fetch_ucirepo(id = 274)
sml2010_data <- sml2010$data$original

dim(sml2010_data)
head(sml2010_data)
```

- **Number of features:** not clearly specified on the UCI page
- **Number of targets:** not clearly specified on the UCI page
- **Number of observations:** 4,137
- **Response variable:** a natural regression target is an indoor temperature variable such as the dining-room or room temperature
- **Topic:** smart-home environmental sensor measurements
- **Missing-data handling:** inspect and handle possible missing values; the UCI page says the dataset could contain missing values
- **Parsing / encoding:** parse the UTC `Date` and `Time` fields and skip the commented header lines that begin with `#`
- **Merging / reshaping:** combine `NEW-DATA-1.T15.txt` and `NEW-DATA-2.T15.txt` into one dataset -->

<!-- ## Online Retail II

Webpage: <https://archive.ics.uci.edu/dataset/502/online+retail+ii>

```{r}
#| eval: false
online_retail_ii <- fetch_ucirepo(id = 502)
online_retail_ii_data <- online_retail_ii$data$original

dim(online_retail_ii_data)
head(online_retail_ii_data)
```

- **Number of features:** not clearly specified on the UCI page
- **Number of targets:** not clearly specified on the UCI page
- **Number of observations:** 1,067,371
- **Response variable:** a workable quantitative target is `UnitPrice` or an aggregated revenue outcome
- **Topic:** multi-year online retail transactions
- **Missing-data handling:** inspect and handle missing values, especially in customer-level fields
- **Parsing / encoding:** parse `InvoiceDate` as a date-time variable
- **Merging / reshaping:** combine the workbook sheets into one consistent transaction table before modeling -->

<!-- ## Open University Learning Analytics Dataset

Webpage: <https://archive.ics.uci.edu/dataset/349/open+university+learning+analytics+dataset>

```{r}
#| eval: false
oulad <- fetch_ucirepo(id = 349)
oulad_data <- oulad$data$original

dim(oulad_data)
head(oulad_data)
```

- **Number of features:** not clearly specified on the UCI page
- **Number of targets:** not clearly specified on the UCI page
- **Number of observations:** not clearly specified on the UCI page
- **Response variable:** a workable quantitative target is assessment `score` after joining the student and assessment tables
- **Topic:** student activity, registration, and assessment behavior in online courses
- **Missing-data handling:** inspect and handle missing values; the UCI page marks the dataset as having missing values
- **Parsing / encoding:** convert date-like offset fields such as registration day, assessment day, and VLE interaction day into a consistent course-time scale
- **Merging / reshaping:** join the linked tables such as `studentAssessment.csv`, `studentInfo.csv`, `studentRegistration.csv`, `studentVle.csv`, and `vle.csv` -->

<!-- ## Restaurant & consumer data

Webpage: <https://archive.ics.uci.edu/dataset/232/restaurant+consumer+data>

```{r}
#| eval: false
restaurant_consumer <- fetch_ucirepo(id = 232)
restaurant_consumer_data <- restaurant_consumer$data$original

dim(restaurant_consumer_data)
head(restaurant_consumer_data)
```

- **Number of features:** not clearly specified on the UCI page
- **Number of targets:** not clearly specified on the UCI page
- **Number of observations:** the main rating table has 1,161 user-place ratings; the auxiliary tables range from 130 to 2,339 rows
- **Response variable:** a clear regression target is `rating` from `rating_final.csv`
- **Topic:** restaurant recommendation with user, place, and context attributes
- **Missing-data handling:** inspect and handle missing values in files such as `geoplaces2.csv` and `userprofile.csv`
- **Parsing / encoding:** encode the many categorical preference and restaurant-description variables
- **Merging / reshaping:** join `rating_final.csv` with the user and restaurant tables using `userID` and `placeID` -->

# Final remark

A few of the datasets above are more naturally classification datasets, but they were included because they were in your earlier long list. If you want, I can also make a **strict version** that keeps only datasets that clearly satisfy all of your regression-project rules without any caveats.
