## Communities and Crime Unnormalized

Webpage: <https://archive.ics.uci.edu/dataset/211/communities+and+crime+unnormalized>

```{r}
communities_crime <- fetch_ucirepo(id = 211)
communities_crime_data <- communities_crime$data$original

dim(communities_crime_data)
head(communities_crime_data)
```

- **Number of features:** 125
- **Number of targets:** 18 possible crime outcomes are described on the UCI page
- **Number of observations:** 2,215
- **Response variable:** a clear regression target is `ViolentCrimesPerPop`
- **Topic:** socio-economic, law-enforcement, and crime measurements for U.S. communities
- **Missing-data handling:** inspect and handle the many missing values, especially in the law-enforcement variables
- **Parsing / encoding:** encode categorical identifiers such as `State` if they are retained
- **Other preparation:** choose one crime variable as the target and remove overlapping crime subtotal variables that would leak information into that target

<!-- ## KDD Cup 1998

Webpage: <https://archive.ics.uci.edu/dataset/129/kdd+cup+1998>

```{r}
#| eval: false
kdd_cup_1998 <- fetch_ucirepo(id = 129)
kdd_cup_1998_data <- kdd_cup_1998$data$original

dim(kdd_cup_1998_data)
head(kdd_cup_1998_data)
```

- **Number of features:** not clearly specified on the UCI page
- **Number of targets:** not clearly specified on the UCI page
- **Number of observations:** 191,779
- **Response variable:** a clear regression target is `TARGET_D`, the donation amount in the original KDD target files
- **Topic:** direct-marketing donation modeling
- **Missing-data handling:** inspect and handle the missing values noted on the UCI page
- **Parsing / encoding:** encode the large number of categorical marketing and demographic variables
- **Merging / reshaping:** join the learning file with the provided target file information when reconstructing a modeling table -->

