# Adult income dataset
This dataset is from [UCI machine learning repository](https://archive.ics.uci.edu/dataset/2/adult). The dataset is about the annual income of an individual and the purpose of the project is to predict whether the annual income exceeds $50K/year or not. 

## 







## Reading data
- There are two dataset files. `adult.data` is the training dataset and `adult.test` is the test dataset.
- Both files don't have header lines. So we need to manually specify column names. 
- The first line of `adult.test` should be skipped since it is irrelevent to our dataset.
- In this dataset missing values are marked as ` ?`. We would like to remove all missing values.


```{python}
#| eval: false
import pandas as pd

train_df = pd.read_csv("../assests/datasets/adult/adult.data", na_values=" ?", header=None).dropna()
test_df = pd.read_csv("../assests/datasets/adult/adult.test", na_values=" ?", skiprows=[0], header=None).dropna()

colnames = [
    "age",
    "workclass",
    "fnlwgt",
    "education",
    "education-num",
    "marital-status",
    "occupation",
    "relationship",
    "race",
    "sex",
    "capital-gain",
    "capital-loss",
    "hours-per-week",
    "native-country",
    "income",
]

train_df.columns = colnames
test_df.columns = colnames
```

## Preprocessing
There are 15 columns. The last column is `income` which has two string values `' >50K'` and `' <=50K'`. The rest variables are descibed below.

- `age`: continuous
- `workclass`: 8 categories
- `fnlwgt`: continuous
- `education`: 16 categories
- `education-num`: continuous
- `marital-status`: 7 categories
- `occupation`: 14 categories
- `relationship`: 6 categories
- `race`: 5 categories
- `sex`: 2 categories
- `capital-gain`: continuous
- `capital-loss`: continuous
- `hours-per-week`: continuous
- `native-country`: 41 categories

So for features we need to normalize numerical columns and apply one-hot encoding for categorical columns. For labels, we need to convert `' >50K'` and `' <=50K'` into `1` and `0`.

### Split the label column


```{python}
#| eval: false
X_train = train_df.drop("income", axis=1)
y_train = (train_df["income"] == " >50K").astype(int)
X_test = test_df.drop("income", axis=1) 
y_test = (test_df["income"] == " >50K").astype(int)
```



### Normalize
```{python}
#| eval: false
from sklearn.preprocessing import MinMaxScaler

numeric_features = [
    "age",
    "fnlwgt",
    "education-num",
    "capital-gain",
    "capital-loss",
    "hours-per-week",
]
mm = MinMaxScaler()

numeric_X_train = mm.fit_transform(X_train[numeric_features])
numeric_X_test = mm.transform(X_test[numeric_features])
```

### One-hot encoding

```{python}
#| eval: false
from sklearn.preprocessing import OneHotEncoder

categorical_features = [
    "workclass",
    "education",
    "marital-status",
    "occupation",
    "relationship",
    "race",
    "sex",
    "native-country",
]
ohe = OneHotEncoder(handle_unknown="ignore")

categorical_X_train = ohe.fit_transform(X_train[categorical_features])
categorical_X_test = ohe.transform(X_test[categorical_features])
```

The onehot decoding of the categorical columns is relative large, and is therefore stored as a sparse matrix (which is an object different from numpy array). So we need to convert it back to numpy array.

```{python}
#| eval: false
import numpy as np

categorical_X_train = categorical_X_train.toarray()
categorical_X_test = categorical_X_test.toarray()
```


### Concatenate
```{python}
#| eval: false
X_train_norm = np.concatenate([numeric_X_train, categorical_X_train], axis=1)
X_test_norm = np.concatenate([numeric_X_test, categorical_X_test], axis=1)
```

### Composer

The best pratice for the above operations is to write a composer.


```{python}
#| eval: false
from sklearn.compose import ColumnTransformer
from sklearn.preprocessing import StandardScaler, OneHotEncoder

preprocess = ColumnTransformer(
    transformers=[
        ("num", StandardScaler(), numeric_features),
        ("cat", OneHotEncoder(handle_unknown="ignore"), categorical_features),
    ]
)

X_train_norm = preprocess.fit_transform(X_train)
X_test_norm = preprocess.transform(X_test)
```

The results here are also sparse matrix. So we need to convert it back to dense array.


```{python}
#| eval: false
X_train_norm = X_train_norm.toarray()
X_test_norm = X_test_norm.toarray()
```




## `Dataset` class

We build a `Dataset` class in order to use it for a nerual network.


```{python}
#| eval: false
import torch
from torch.utils.data import Dataset

class AdultDataset(Dataset):
    def __init__(self, X, y):
        self.X = torch.tensor(X, dtype=torch.float32)
        self.y = torch.tensor(y, dtype=torch.long).reshape(-1, 1)


    def __len__(self):
        return len(self.X)

    def __getitem__(self, idx):
        return self.X[idx], self.y[idx]


train_ds = AdultDataset(X_train_norm, y_train)
test_ds = AdultDataset(X_test_norm, y_test)
```


## Regression Variant

Webpage: <https://archive.ics.uci.edu/dataset/2/adult>

```{r}
adult <- fetch_ucirepo(id = 2)
adult_data <- adult$data$original

dim(adult_data)
head(adult_data)
```

- **Number of features:** 14
- **Number of targets:** 1
- **Number of observations:** 48,842
- **Response variable:** if you want regression, a workable quantitative outcome is `hours-per-week`
- **Topic:** census-style demographic and employment data
- **Missing-data handling:** convert `?` placeholders to `NA`
- **Parsing / encoding:** one-hot encode categorical predictors such as work class, education, marital status, occupation, relationship, race, sex, and native country
- **Other preparation:** remove or redefine the original classification target if it is not part of the regression task

<!-- ## Bank Marketing

Webpage: <https://archive.ics.uci.edu/dataset/222/bank+marketing>

```{r}
#| eval: false
bank_marketing <- fetch_ucirepo(id = 222)
bank_marketing_data <- bank_marketing$data$original

dim(bank_marketing_data)
head(bank_marketing_data)
```

- **Number of features:** 16
- **Number of targets:** 1
- **Number of observations:** 45,211
- **Response variable:** official target is `y`; for regression, users sometimes instead model a quantitative field such as `balance`
- **Topic:** direct marketing campaigns by a Portuguese banking institution
- **Preprocessing needed:** remove or carefully handle the campaign outcome variable `y` if it is not being modeled
- **Preprocessing needed:** encode categorical fields such as job, marital status, education, contact type, month, and poutcome
- **Preprocessing needed:** check for skewness in `balance`, `duration`, and campaign-count variables and transform if needed -->

<!-- ## Online Retail

Webpage: <https://archive.ics.uci.edu/dataset/352/online+retail>

```{r}
#| eval: false
online_retail <- fetch_ucirepo(id = 352)
online_retail_data <- online_retail$data$original

dim(online_retail_data)
head(online_retail_data)
```

- **Number of variables:** 8
- **Number of observations:** 541,909
- **Response variable:** a quantitative choice is `UnitPrice`
- **Topic:** online retail transactions
- **Preprocessing needed:** parse `InvoiceDate` as a date-time variable
- **Preprocessing needed:** remove cancellations and returns, typically identified by invoice codes beginning with `C` or by negative quantities
- **Preprocessing needed:** decide how to handle missing `CustomerID`
- **Preprocessing needed:** encode `Country` and engineer transaction-level or customer-level features if modeling beyond raw line items -->

<!-- ## Car Evaluation

Webpage: <https://archive.ics.uci.edu/dataset/19/car+evaluation>

```{r}
#| eval: false
car_evaluation <- fetch_ucirepo(id = 19)
car_evaluation_data <- car_evaluation$data$original

dim(car_evaluation_data)
head(car_evaluation_data)
```

- **Number of features:** 6
- **Number of targets:** 1
- **Number of observations:** 1,728
- **Response variable:** this is mainly a classification dataset; if you use it for regression, you would need to create an ordinal or numeric response yourself
- **Topic:** car acceptability evaluation
- **Preprocessing needed:** create a numeric or ordinal response from the categorical acceptability label
- **Preprocessing needed:** encode all predictors, which are categorical and ordinal
- **Preprocessing needed:** preserve ordinal structure for variables such as buying price, maintenance cost, safety, luggage boot size, and number of persons -->
